Our Title

### 2015-05-01 draft.tex
Corrected US/U.S. notation.

Where appropriate, expected countermeasures, such as shielding of radioactive sources, are discussed within the context of discerning a weak signal in a strong background. In circumstances where it is deemed essential for personnel to monitor the shipment of regular commerce or travel, including but not limited to commercial or private traffic by land, sea, or air, portal detectors and active interrogation is considered. Such locations are qualified for observation which are considered high risk or of strategic value in fielding a conventional response to attack from ISIL on the US, allies, or any sovereign territory for which instability is a matter of US National Security. It is assumed that ISIL will prioritize countervalue over counterforce targets, since countervalue targeting is more disruptive and more effective in a campaign of terror.

@@

    Throughout the many conflicts in Iraq in the remainder of the 2000s, AQI underwent many changes in leadership, ideology, and name. These changes, primarily the deaths of much of its upper leadership (including al-Zarqawi), were caused by a combination of military pressure from US troops, multiple disagreements  with bin Laden's al-Qaeda, and backlash from other Sunnis in Iraq \cite{Zelin2014,Kahl2008}. As a result of this, AQI was reorganized as the Islamic State of Iraq (ISI) in 2007, with Abu Bakr al-Baghdadi as it newly appointed leader \cite{Zelin2014,Shadid2010}. Under the new leadership of al-Baghdadi, ISI broke a period of regional detente with al-Qaeda (following the death of Osama bin Laden in August 2011), driving a further wedge between the two groups by al-Baghdadi's refusal to pledge allegiance to the new leader of al-Qaeda, Ayman al-Zawahiri \cite{Al-Jawlani}. In April 2013, this division resulted in  al-Baghdadi announcing that ISI would be extending its reach into Syria, and changing its name to the Islamic State of Iraq and al-Sham, or the Islamic State of Iraq and the Levant \cite{Zelin2014,Al-Hussaini2013}. In June 2014, al-Baghdadi declared ISIL a global Caliphate, with himself as its Caliph \cite{Mortada2014,TheWeek2014}. To understand the significance of this declaration, let us consider the history and role of Caliphs and Caliphates in Islam.

@@

ISIL has many enemies, indeed many more than are graphically depicted in \autoref{fig:ISILRelationships}.  As of January 2015, the anti-ISIL coalition led by the US included 62 nations and groups including the US, Canada, Iraq, Jordan, Bahrain, Saudi Arabia, UAE, France, Germany, United Kingdom, Australia, Belgium, Denmark, Italy, Czech Republic, Albania, Netherlands, Estonia, Hungary, Turkey, and Lebanon, all of whom have carried out or supported military operations against ISIL \cite{Wordsworth2015}.  Additionally, although not allied with the US coalition, Syria, Iran, and various Kurdish, Shia, and Sunni militias in Syria and Iraq have opposed ISIL \cite{Mooney2014}.  Despite this broad coalition, there are many unlikely allies in the list which present potential seams for ISIL exploitation  through propaganda or military action.  For example, strikes against ISIL in Syria benefit the Assad regime while US trained rebels could turn on and attack the Assad regime \cite{Shinkman2015}.  Officially, the US supports the removal of the Assad regime albeit without direct military intervention on part of the US \cite{Shinkman2015}.  Similarly, US airstrikes and intelligence have supported Iran and Iraqi Sunni and Shia militias (including several that once fought US forces) in their campaigns against ISIL \cite{Chulov2014}.  While these uneasy alliances have placed pressure on ISIL, it is unclear if they can be sustained over the timeframe that would be necessary to defeat ISIL and prevent their acquisition of a nuclear weapon. Sustaining these alliances would be especially difficult if ISIL were to turn its formidable propaganda machine to breaking some of these fragile links.

Additionally, there are longstanding mistrust, perceptions, and illicit alliances that could cause large fissures  without the help of ISIL.  While the Iraqi government is cooperating with the US in battling ISIL, some clerics and government officials have publicly claimed the CIA was responsible for the rise of ISIL \cite{Kirkpatrick2014}.  Whether this is true is largely irrelevant if the mere perception affects the extent of US - Iraqi cooperation against ISIL, especially if the need would arise to place US troops in combat.  Other reports have the Turkey military providing safe travel and weapons for ISIL fighters attacking Kurdish forces, which have been a thorn in the side of Turkey \cite{Guiton2014}.  Finally, it appears that wealthy Saudis, who have a history of funding terrorist organizations, are continuing to provide ISIL with funding \cite{Windrem2014}. While this funding and free travel is not a large contributor to ISIL’s overall success, it represents larger issues faced in isolating ISIL from the international community and shows a willingness of certain groups to deal with ISIL if they see personal benefit despite the international condemnation.  Currently, it is just guns, cash, and ammunition, but as time wears on does someone sell nuclear components or technology to ISIL in an attempt to accomplish some goal?  It may sound unfathomable, but perhaps no more irrational that the US arming militias and rebels in Syria, despite the fact that the current issues facing the Middle East with regards to terrorism and ISIL can be traced to the US arming and training of Afghani rebels in the 1980s.

@@

Finally, it is important to consider the non-organizationally alligned support that ISIS recieves in the form of foriegn fighters.  The flow of these foriegn fighters has dwarfed the influx associated with previous calls to Jihad, such as Afghanistan in the 1980s \cite{Barrett2014}.  These fighters come from all over the world as shown in \autoref{fig:enablers}, and they bring a diverse background and skill set with them.  Although no comprehensive list has been established in the open literature about the skills brought to ISIL via foreign fighters, it is likely that some of the foreign fighters possess skills necessarily to develop, fabricate, and assemble components of a nuclear weapon.

@@

The threat of  nuclear weapon states is not a novel thought. Being the first and only nation to use a nuclear weapon against an enemy, the United States has feared a nuclear threat ever since weaponized nuclear technology became possible. Shortly after, the Soviet Union acquired nuclear weapons, marking the onset of the nuclear standoff during the Cold War.   Fortunately, through deterrence based on game theory against a rational adversary, not a single shot was fired between the US and the USSR for approximately 44 years, until the conflict was alleviated. Since then, the nuclear threat has spread to many nations, including, most recently, Iran.

@@

This  report will evaluate and discuss the ability of  ISIL to engage in nuclear terrorism, so as to inform a policy statement to be enforced by the US. This report considers the capability and resources of other sovereign powers and terrorist organizations when those may benefit ISIL. This policy statement details an organized response to any step ISIL may make towards the deployment of a device containing special nuclear material (SNM). It stops short of a larger policy directed towards  the Administration's goal to degrade and ultimately destroy ISIL.

@@

The most substantial effect of an RDD is the spread of fear, panic, economic disruption, and the cost associated with cleaning up a contaminated area. These devices are unlikely to result in as many deaths as a fission or fusion device due to the short duration of radiation exposure and the diffuse distribution of the dispersed materials. \autoref{tab:health_effects} from the US Environmental Protection Agency shows the level of exposure required to produce various health symptoms and the time from acute exposure to symptom onset \cite{USEPA1999}. Department of Homeland Security (DHS) and local law enforcement crisis management plans to evacuate crowds following an RDD detonation and the subsequent radiation cleanup will drastically reduce the number of people who receive radiation doses large enough to cause short term health effects.

@@

The situation is only complicated when considering international intelligence agencies and the sharing of data. To simplify the problem, the factors considered in determining the level of cooperation and the nature of an intelligence partnership between US and other countries were distilled down to the historical relationships between both parties, the country’s relationship to ISIL, its military and intelligence capability, and its vulnerability and geographical proximity to ISIL. The role of international entities is gauged by the organization’s functions, credibility, and past success in relaying information and promptly responding to crises.  These factors are further described in \autoref{app:intel}.

@@

\begin{itemize}
  \item \textbf{Level 1:} US should install open information sharing with the countries in category one. During WWII, US and Great Britain had joint research of the atomic bomb. This cooperation can be repeated for actions against ISIL.
  \item \textbf{Level 2:} Information sharing with category two countries should be limited to low and medium security information (see \autoref{app:intel} for further description).  
  \item \textbf{Level 3:} The U.S. should consider to have limited information sharing of low importance security information with category three countries.  
\end{itemize}

@@

UK and France have been US’ allies even since the country launched its global fight against terrorism after the 9/11 incident. Hence, the US can expect UK specifically to be its closest ally in preventing ISIL’ nuclear terrorism effort. The two countries \enquote{special relationship} dated back to Churchillian era means that the US can usually count on the UK for mutual support, especially considering ISIL’ threats to the country is very high. UK has been ISIL’ main target for terrorist attacks, and given UK’s nuclear capability, the country is extra vulnerable to terrorist attacks on any of its nuclear installations. So far, UK has joined US-led coalition against ISIL, nevertheless, prospect for bilateral cooperation between the two countries cannot be disregarded.

Meanwhile, France is US’ oldest ally and relationship remains active and friendly. France supported US in various issues, however, France was skeptical when US invaded Iraq after the 9/11 incident. ISIL has also posed great threats to France (being a successful nuclear energy producer and all), France can be vulnerable to attacks on its nuclear energy facilities as well as its possession of nuclear weapons stocks. France is also among ISIL’ main recruitments considering France high marginalized and discontented Muslim populations. France’s high ISIL threats and vulnerabilities made the country a good collaborator in effectively fighting ISIL.

Nuclear capabilities in the hand of weak or unstable government give opportunity for ISIL to acquire nuclear capabilities. Nuclear capabilities of Pakistan and its weak governance as well as security vulnerabilities is of concern to the US. Looking at US-Pakistan relationship, it has its ups and down. Over the decades, Pakistan has been a great ally to the US, but there are a few instance where the two countries were in disagreement. US viewed Pakistan as its ally and threat in the country’s global war on terror, in which Pakistan is seen as the ‘weakest link’ as it continue to fail in containing act of terrorism going on in its own country. US need to prop Pakistan and reduce the country’s vulnerabilities of ISIL penetrations and the possibility of ISIL acquiring Pakistan’s nuclear weapons.

India is also another country with nuclear weapon capability, and is mostly on the \enquote{other side} of US’ security policies. India, together with Russia was among the pioneers of Non-Aligned Movement (NAM), in which views on security issues tend to be divergent with the US. However, as of recently, the US and India have set up ‘global strategic partnership’, based on shared democratic values and increasing convergence of interests on bilateral, regional and global issues; key recent developments in the relationship include the rapid growth of India's economy and bilateral trade, the close links between the Indian and American computer and Internet industries, a geopolitical coalition to balance the rise of China, the weakening of US-Pakistan relations, among others. US has cooperated with India in area of civil nuclear, and this can be expanded to development of nuclear security hopefully in the future. However, unlike Pakistan, India believed ISIL’ threats as periphery issue to the country.

In analyzing Russia as a possible bilateral partner, the two country has had strained relationship, increasingly since the economic sanctions imposed by the US to Russia. Dispute between the two countries have centered on power struggle for new world order, and recently exacerbated by Russia’s annexation of Crimea. The two countries have been historical rival and the Cold War has proven to be one of the expansive and stretched out power struggle between the two. However, both countries realized and agreed to maintain dialogue in several areas including terrorism. Being misaligned with Russia on ISIL and nuclear terrorism issue could be serious mistakes as Russia’s nuclear capability, if not protected and are allowed to be shared in the wrong hand, could have grave catastrophic impact to global security environment.

China, another US’ partner and rival at times, also have nuclear capabilities. China and the US has close partnership particularly in economic sector, two-way trade between the two countries is over US\$590 billion in 2014 alone. Chinese Muslim Uyghur militants have been associated with ISIL and efforts to undertake terrorist attacks in China. This strengthen the need for collaboration between China and US; to ensure any ISIL activities can be contained in China and the country’s nuclear weapons is protected from any probability of ISIL acquiring them.

US considered North Korea as grave threats and this can be further exacerbated if North Korea is presented with the opportunity to work with ISIL. The prospect of North Korea selling their nuclear weapons and missiles to ISIL cannot be discounted. However, the possibility of US and North Korea to cooperate in hampering this issue is not viable. The US need to come up with initiatives to solve this problem.

Iran and US shared a common enemy factor in terms of ISIL. As the nuclear talks between the two countries are being hashed out, both countries have committed on fighting ISIL together. This can be a good opportunity for US to combat ISIL more effectively as US cannot disregard Iran’s influence and role in Iraq’s security landscape and their fight against ISIL.
Meanwhile, collaborating with Israel is considered a gray area and not clear-cut as it seems. Israel has been known to voice their support of ISIL, keeping in line with ISIL’s common enemy with Israel i.e. Iran. As US and Israel have been a strong ally in the Middle East region, this fact presented a conundrum to the US.

The US may find other allies in the Middle East region, however, the region is deeply divided in their positions against the ISIL. Although most Middle Eastern countries sees ISIL as a serious threat to the region and its people, some still supported ISIL, if not openly but explicitly. US has worked closely with Jordan, with the US creating a US\$5 billion counterterrorism partnership fund with Jordan and other countries (Libya, Lebanon, Syria) that are considered on the \enquote{frontlines} of US’ fights against ISIL.

The US can find allies in European countries as well, as the region is facing \enquote{recruitment} threats from ISIL. Thousands of Europeans have been recruited by ISIL to fight jihad cause in ISIL’ territories. ISIL is also setting up bases in a number of European countries, expanding its influence beyond the Middle East region. The US can identify these countries and create cooperation efforts in order to effectively eliminate ISIL.

@@

As ISIL carving its traction in Iraq and Syria, the two countries become an immediate concern for help by the US. Strengthening Iraq and Syria could be among the first line of defense against ISIL. As US has an improving bilateral relations with Iraq, Iraq become an easier partner for US to work with in reigning ISIL. The Strategic Framework Agreement for a Relationship of Friendship and Cooperation between the US and the Republic of Iraq guides the overall political, economic, cultural, and security ties with Iraq.  This agreement is designed to help the Iraqi people stand on their own and reinforce Iraqi sovereignty, while protecting US interests in the Middle East.

However, the country’s relation with Syria is not black-and-white as we would like it to be. The US-Syria relationship is negative after having suspended diplomatic relations in 2014. The US has also imposed economic sanction to Syria since 2004 due to the country’s involvement in harboring terrorist groups and the meddling of Lebanon’s political affairs including the assassination of Lebanese Prime Minister Rafik Hariri. The US deeply opposed to the ongoing Syrian Civil war and President Bashar al-Assad’s regime, specifically regarding the country’s use and stockpile of chemical and biological weapons. The US has floated the idea to train the Syrian military opposition, and even considering to train the Syrian military itself in order to fight ISIL, nevertheless, President Assad viewed the proposal as infringing on their country’s sovereignty. Despite the two countries differences, they actually both share a common enemy, and it is hope this fact can be harnessed to US’ advantage.

@@

Current US strategy is to:

@@

In the current oil market, diminished profits from selling oil may make ISIL more dependent on external donations. As need for this arises, it is important to focus on these external sources. Saudi Arabia and UAE are becoming more effective at preventing funding of terrorist orgnaizations by their citizens, but Kuwait and Qatar remain permissive jurisdictions. More needs to be done by these states to enforce their newly passed laws against financing terrorism, and the US can help them do this through State Department and other training. It is important to note that these Gulf states have an interest in stopping ISIL – ISIL has periodically targeted Qatar and its Ministry of Foreign Affairs. \subsubsection{???????? THIS PARAGRAPH NEEDS SOURCES   ?????????}

@@

Disabling ISIL oil revenue is more complicated than it might seem. ISIL has been focused on establishing a durable internal market for oil, ensuring they have a reliable source of fuel for its fleets and also establishing dependence from the local population on its cheap oil. Recent airstrikes have targeted the oil at its source, but might be better served by targeting means of transportation and sale. Syrian rebels have pointed to this as a major error in targeting, as it disrupts oil supply to power generators and local facilities, allowing ISIL to pass the blame on to the US (who they call \enquote{the crusaders}) when the electricity system failes. One difficulty in targeting the means of transportation and sale is that oil is transported in tanker trucks and oil drums, and most of the transactions are in cash, making them hard to track. One policy priority should be providing large quantities of diesel fuel and oil for generators in opposition areas of northern Syria [1].
A final recommendation is to put more intelligence resources toward detecting ISIL fundraising efforts through modern communication networks (social media) \cite{Report2015}.

@@

The US should share practical information and intelligence at an international level, both spontaneously and on request, to effectively disrupt international financial flows \cite{Report2015}.

@@

The US should request countries to proactively identify individuals and entities for inclusion in the UN Al Qaida Sanctions Committee list \cite{Report2015}.

@@

ISIL has shown itself prolific at recruiting foreign fighters, and the US must try to stem this flow in its entireity, but alos to prevent nuclear scientists from joining ISIL's causes.

@@

ISIL could attempt to purchase or steal a nuclear weapon, and so the nations which maintain a stockpile of nuclear weapons must remain under scrutiny. Nations of the highest interest that have nuclear weapons are Pakistan, North Korea, India, Turkey, Israel, and Russia. Israel has not been confirmed to have nuclear weapons, but it is widely believed that it has nuclear capabilites. Turkey is a NATO nuclear weapon sharing state and has US tactical nuclear weapons stationed within its borders. However, it is unlikely that ISIL would be able to steal a nuclear weapon from either of these states, as those in Turkey are under US protection and Israel has proven adept at hiding their arsenal. Of greatest concern is the possibility that any of these nuclear-capable states, North Korea in particular, would willingly sell a physics package to ISIL.

@@

The US sees merits in pursuing both multilateral and bilateral effort in international strategic communications. Most of the times, the US tends to favored intense bilateralism over multilateralism to achieve its objectives. However, following World War II, multilateral initiatives in the area of nuclear non-proliferation and disarmament increased in significance.  This multilateral transition was marked by the creation of the United Nations (UN) in 1945.  The UN, and several subsequent agencies, are the current approach taken by the U.S. to address nuclear proliferation concerns.

>>>>

Where appropriate, expected countermeasures, such as shielding of radioactive sources, are discussed within the context of discerning a weak signal in a strong background. In circumstances where it is deemed essential for personnel to monitor the shipment of regular commerce or travel, including but not limited to commercial or private traffic by land, sea, or air, portal detectors and active interrogation is considered. Such locations are qualified for observation which are considered high risk or of strategic value in fielding a conventional response to attack from ISIL on the U.S., allies, or any sovereign territory for which instability is a matter of U.S. National Security. It is assumed that ISIL will prioritize countervalue over counterforce targets, since countervalue targeting is more disruptive and more effective in a campaign of terror.

@@

    Throughout the many conflicts in Iraq in the remainder of the 2000s, AQI underwent many changes in leadership, ideology, and name. These changes, primarily the deaths of much of its upper leadership (including al-Zarqawi), were caused by a combination of military pressure from U.S. troops, multiple disagreements  with bin Laden's al-Qaeda, and backlash from other Sunnis in Iraq \cite{Zelin2014,Kahl2008}. As a result of this, AQI was reorganized as the Islamic State of Iraq (ISI) in 2007, with Abu Bakr al-Baghdadi as it newly appointed leader \cite{Zelin2014,Shadid2010}. Under the new leadership of al-Baghdadi, ISI broke a period of regional detente with al-Qaeda (following the death of Osama bin Laden in August 2011), driving a further wedge between the two groups by al-Baghdadi's refusal to pledge allegiance to the new leader of al-Qaeda, Ayman al-Zawahiri \cite{Al-Jawlani}. In April 2013, this division resulted in  al-Baghdadi announcing that ISI would be extending its reach into Syria, and changing its name to the Islamic State of Iraq and al-Sham, or the Islamic State of Iraq and the Levant \cite{Zelin2014,Al-Hussaini2013}. In June 2014, al-Baghdadi declared ISIL a global Caliphate, with himself as its Caliph \cite{Mortada2014,TheWeek2014}. To understand the significance of this declaration, let us consider the history and role of Caliphs and Caliphates in Islam.

@@

ISIL has many enemies, indeed many more than are graphically depicted in \autoref{fig:ISILRelationships}.  As of January 2015, the anti-ISIL coalition led by the U.S. included 62 nations and groups including the U.S., Canada, Iraq, Jordan, Bahrain, Saudi Arabia, UAE, France, Germany, United Kingdom, Australia, Belgium, Denmark, Italy, Czech Republic, Albania, Netherlands, Estonia, Hungary, Turkey, and Lebanon, all of whom have carried out or supported military operations against ISIL \cite{Wordsworth2015}.  Additionally, although not allied with the U.S. coalition, Syria, Iran, and various Kurdish, Shia, and Sunni militias in Syria and Iraq have opposed ISIL \cite{Mooney2014}.  Despite this broad coalition, there are many unlikely allies in the list which present potential seams for ISIL exploitation  through propaganda or military action.  For example, strikes against ISIL in Syria benefit the Assad regime while U.S. trained rebels could turn on and attack the Assad regime \cite{Shinkman2015}.  Officially, the U.S. supports the removal of the Assad regime albeit without direct military intervention on part of the U.S. \cite{Shinkman2015}.  Similarly, U.S. airstrikes and intelligence have supported Iran and Iraqi Sunni and Shia militias (including several that once fought U.S. forces) in their campaigns against ISIL \cite{Chulov2014}.  While these uneasy alliances have placed pressure on ISIL, it is unclear if they can be sustained over the timeframe that would be necessary to defeat ISIL and prevent their acquisition of a nuclear weapon. Sustaining these alliances would be especially difficult if ISIL were to turn its formidable propaganda machine to breaking some of these fragile links.

Additionally, there are longstanding mistrust, perceptions, and illicit alliances that could cause large fissures  without the help of ISIL.  While the Iraqi government is cooperating with the U.S. in battling ISIL, some clerics and government officials have publicly claimed the CIA was responsible for the rise of ISIL \cite{Kirkpatrick2014}.  Whether this is true is largely irrelevant if the mere perception affects the extent of U.S. - Iraqi cooperation against ISIL, especially if the need would arise to place U.S. troops in combat.  Other reports have the Turkey military providing safe travel and weapons for ISIL fighters attacking Kurdish forces, which have been a thorn in the side of Turkey \cite{Guiton2014}.  Finally, it appears that wealthy Saudis, who have a history of funding terrorist organizations, are continuing to provide ISIL with funding \cite{Windrem2014}. While this funding and free travel is not a large contributor to ISIL’s overall success, it represents larger issues faced in isolating ISIL from the international community and shows a willingness of certain groups to deal with ISIL if they see personal benefit despite the international condemnation.  Currently, it is just guns, cash, and ammunition, but as time wears on does someone sell nuclear components or technology to ISIL in an attempt to accomplish some goal?  It may sound unfathomable, but perhaps no more irrational that the U.S. arming militias and rebels in Syria, despite the fact that the current issues facing the Middle East with regards to terrorism and ISIL can be traced to the U.S. arming and training of Afghani rebels in the 1980s.

@@

Finally, it is important to consider the non-organizationally aligned support that ISIL receives in the form of foreign fighters.  The flow of these foreign fighters has dwarfed the influx associated with previous calls to Jihad, such as Afghanistan in the 1980s \cite{Barrett2014}.  These fighters come from all over the world as shown in \autoref{fig:enablers}, and they bring a diverse background and skill set with them.  Although no comprehensive list has been established in the open literature about the skills brought to ISIL via foreign fighters, it is likely that some of the foreign fighters possess skills necessarily to develop, fabricate, and assemble components of a nuclear weapon.

@@

The threat of  nuclear weapon states is not a novel thought. Being the first and only nation to use a nuclear weapon against an enemy, the United States has feared a nuclear threat ever since weaponized nuclear technology became possible. Shortly after, the Soviet Union acquired nuclear weapons, marking the onset of the nuclear standoff during the Cold War.   Fortunately, through deterrence based on game theory against a rational adversary, not a single shot was fired between the U.S. and the USSR for approximately 44 years, until the conflict was alleviated. Since then, the nuclear threat has spread to many nations, including, most recently, Iran.

@@

This  report will evaluate and discuss the ability of  ISIL to engage in nuclear terrorism, so as to inform a policy statement to be enforced by the U.S.. This report considers the capability and resources of other sovereign powers and terrorist organizations when those may benefit ISIL. This policy statement details an organized response to any step ISIL may make towards the deployment of a device containing special nuclear material (SNM). It stops short of a larger policy directed towards  the Administration's goal to degrade and ultimately destroy ISIL.

@@

The most substantial effect of an RDD is the spread of fear, panic, economic disruption, and the cost associated with cleaning up a contaminated area. These devices are unlikely to result in as many deaths as a fission or fusion device due to the short duration of radiation exposure and the diffuse distribution of the dispersed materials. \autoref{tab:health_effects} from the U.S. Environmental Protection Agency shows the level of exposure required to produce various health symptoms and the time from acute exposure to symptom onset \cite{USEPA1999}. Department of Homeland Security (DHS) and local law enforcement crisis management plans to evacuate crowds following an RDD detonation and the subsequent radiation cleanup will drastically reduce the number of people who receive radiation doses large enough to cause short term health effects.

@@

The situation is only complicated when considering international intelligence agencies and the sharing of data. To simplify the problem, the factors considered in determining the level of cooperation and the nature of an intelligence partnership between U.S. and other countries were distilled down to the historical relationships between both parties, the country’s relationship to ISIL, its military and intelligence capability, and its vulnerability and geographical proximity to ISIL. The role of international entities is gauged by the organization’s functions, credibility, and past success in relaying information and promptly responding to crises.  These factors are further described in \autoref{app:intel}.

@@

\begin{itemize}
  \item \textbf{Level 1:} U.S. should install open information sharing with the countries in category one. During WWII, U.S. and Great Britain had joint research of the atomic bomb. This cooperation can be repeated for actions against ISIL.
  \item \textbf{Level 2:} Information sharing with category two countries should be limited to low and medium security information (see \autoref{app:intel} for further description).  
  \item \textbf{Level 3:} The U.S. should consider to have limited information sharing of low importance security information with category three countries.  
\end{itemize}

@@

UK and France have been U.S.’ allies even since the country launched its global fight against terrorism after the 9/11 incident. Hence, the U.S. can expect UK specifically to be its closest ally in preventing ISIL’ nuclear terrorism effort. The two countries \enquote{special relationship} dated back to Churchillian era means that the U.S. can usually count on the UK for mutual support, especially considering ISIL’ threats to the country is very high. UK has been ISIL’ main target for terrorist attacks, and given UK’s nuclear capability, the country is extra vulnerable to terrorist attacks on any of its nuclear installations. So far, UK has joined U.S.-led coalition against ISIL, nevertheless, prospect for bilateral cooperation between the two countries cannot be disregarded.

Meanwhile, France is U.S.’ oldest ally and relationship remains active and friendly. France supported U.S. in various issues, however, France was skeptical when U.S. invaded Iraq after the 9/11 incident. ISIL has also posed great threats to France (being a successful nuclear energy producer and all), France can be vulnerable to attacks on its nuclear energy facilities as well as its possession of nuclear weapons stocks. France is also among ISIL’ main recruitments considering France high marginalized and discontented Muslim populations. France’s high ISIL threats and vulnerabilities made the country a good collaborator in effectively fighting ISIL.

Nuclear capabilities in the hand of weak or unstable government give opportunity for ISIL to acquire nuclear capabilities. Nuclear capabilities of Pakistan and its weak governance as well as security vulnerabilities is of concern to the U.S.. Looking at U.S.-Pakistan relationship, it has its ups and down. Over the decades, Pakistan has been a great ally to the U.S., but there are a few instance where the two countries were in disagreement. U.S. viewed Pakistan as its ally and threat in the country’s global war on terror, in which Pakistan is seen as the ‘weakest link’ as it continue to fail in containing act of terrorism going on in its own country. U.S. need to prop Pakistan and reduce the country’s vulnerabilities of ISIL penetrations and the possibility of ISIL acquiring Pakistan’s nuclear weapons.

India is also another country with nuclear weapon capability, and is mostly on the \enquote{other side} of U.S.’ security policies. India, together with Russia was among the pioneers of Non-Aligned Movement (NAM), in which views on security issues tend to be divergent with the U.S.. However, as of recently, the U.S. and India have set up ‘global strategic partnership’, based on shared democratic values and increasing convergence of interests on bilateral, regional and global issues; key recent developments in the relationship include the rapid growth of India's economy and bilateral trade, the close links between the Indian and American computer and Internet industries, a geopolitical coalition to balance the rise of China, the weakening of U.S.-Pakistan relations, among others. U.S. has cooperated with India in area of civil nuclear, and this can be expanded to development of nuclear security hopefully in the future. However, unlike Pakistan, India believed ISIL’ threats as periphery issue to the country.

In analyzing Russia as a possible bilateral partner, the two country has had strained relationship, increasingly since the economic sanctions imposed by the U.S. to Russia. Dispute between the two countries have centered on power struggle for new world order, and recently exacerbated by Russia’s annexation of Crimea. The two countries have been historical rival and the Cold War has proven to be one of the expansive and stretched out power struggle between the two. However, both countries realized and agreed to maintain dialogue in several areas including terrorism. Being misaligned with Russia on ISIL and nuclear terrorism issue could be serious mistakes as Russia’s nuclear capability, if not protected and are allowed to be shared in the wrong hand, could have grave catastrophic impact to global security environment.

China, another U.S.’ partner and rival at times, also have nuclear capabilities. China and the U.S. has close partnership particularly in economic sector, two-way trade between the two countries is over U.S.\$590 billion in 2014 alone. Chinese Muslim Uyghur militants have been associated with ISIL and efforts to undertake terrorist attacks in China. This strengthen the need for collaboration between China and U.S.; to ensure any ISIL activities can be contained in China and the country’s nuclear weapons is protected from any probability of ISIL acquiring them.

U.S. considered North Korea as grave threats and this can be further exacerbated if North Korea is presented with the opportunity to work with ISIL. The prospect of North Korea selling their nuclear weapons and missiles to ISIL cannot be discounted. However, the possibility of U.S. and North Korea to cooperate in hampering this issue is not viable. The U.S. need to come up with initiatives to solve this problem.

Iran and U.S. shared a common enemy factor in terms of ISIL. As the nuclear talks between the two countries are being hashed out, both countries have committed on fighting ISIL together. This can be a good opportunity for U.S. to combat ISIL more effectively as U.S. cannot disregard Iran’s influence and role in Iraq’s security landscape and their fight against ISIL.
Meanwhile, collaborating with Israel is considered a gray area and not clear-cut as it seems. Israel has been known to voice their support of ISIL, keeping in line with ISIL’s common enemy with Israel i.e. Iran. As U.S. and Israel have been a strong ally in the Middle East region, this fact presented a conundrum to the U.S..

The U.S. may find other allies in the Middle East region, however, the region is deeply divided in their positions against the ISIL. Although most Middle Eastern countries sees ISIL as a serious threat to the region and its people, some still supported ISIL, if not openly but explicitly. U.S. has worked closely with Jordan, with the U.S. creating a U.S.\$5 billion counterterrorism partnership fund with Jordan and other countries (Libya, Lebanon, Syria) that are considered on the \enquote{frontlines} of U.S.’ fights against ISIL.

The U.S. can find allies in European countries as well, as the region is facing \enquote{recruitment} threats from ISIL. Thousands of Europeans have been recruited by ISIL to fight jihad cause in ISIL’ territories. ISIL is also setting up bases in a number of European countries, expanding its influence beyond the Middle East region. The U.S. can identify these countries and create cooperation efforts in order to effectively eliminate ISIL.

@@

As ISIL carving its traction in Iraq and Syria, the two countries become an immediate concern for help by the U.S.. Strengthening Iraq and Syria could be among the first line of defense against ISIL. As U.S. has an improving bilateral relations with Iraq, Iraq become an easier partner for U.S. to work with in reigning ISIL. The Strategic Framework Agreement for a Relationship of Friendship and Cooperation between the U.S. and the Republic of Iraq guides the overall political, economic, cultural, and security ties with Iraq.  This agreement is designed to help the Iraqi people stand on their own and reinforce Iraqi sovereignty, while protecting U.S. interests in the Middle East.

However, the country’s relation with Syria is not black-and-white as we would like it to be. The U.S.-Syria relationship is negative after having suspended diplomatic relations in 2014. The U.S. has also imposed economic sanction to Syria since 2004 due to the country’s involvement in harboring terrorist groups and the meddling of Lebanon’s political affairs including the assassination of Lebanese Prime Minister Rafik Hariri. The U.S. deeply opposed to the ongoing Syrian Civil war and President Bashar al-Assad’s regime, specifically regarding the country’s use and stockpile of chemical and biological weapons. The U.S. has floated the idea to train the Syrian military opposition, and even considering to train the Syrian military itself in order to fight ISIL, nevertheless, President Assad viewed the proposal as infringing on their country’s sovereignty. Despite the two countries differences, they actually both share a common enemy, and it is hope this fact can be harnessed to U.S.’ advantage.

@@

Current U.S. strategy is to:

@@

In the current oil market, diminished profits from selling oil may make ISIL more dependent on external donations. As need for this arises, it is important to focus on these external sources. Saudi Arabia and UAE are becoming more effective at preventing funding of terrorist orgnaizations by their citizens, but Kuwait and Qatar remain permissive jurisdictions. More needs to be done by these states to enforce their newly passed laws against financing terrorism, and the U.S. can help them do this through State Department and other training. It is important to note that these Gulf states have an interest in stopping ISIL – ISIL has periodically targeted Qatar and its Ministry of Foreign Affairs. \subsubsection{???????? THIS PARAGRAPH NEEDS SOURCES   ?????????}

@@

Disabling ISIL oil revenue is more complicated than it might seem. ISIL has been focused on establishing a durable internal market for oil, ensuring they have a reliable source of fuel for its fleets and also establishing dependence from the local population on its cheap oil. Recent airstrikes have targeted the oil at its source, but might be better served by targeting means of transportation and sale. Syrian rebels have pointed to this as a major error in targeting, as it disrupts oil supply to power generators and local facilities, allowing ISIL to pass the blame on to the U.S. (who they call \enquote{the crusaders}) when the electricity system failes. One difficulty in targeting the means of transportation and sale is that oil is transported in tanker trucks and oil drums, and most of the transactions are in cash, making them hard to track. One policy priority should be providing large quantities of diesel fuel and oil for generators in opposition areas of northern Syria [1].
A final recommendation is to put more intelligence resources toward detecting ISIL fundraising efforts through modern communication networks (social media) \cite{Report2015}.

@@

The U.S. should share practical information and intelligence at an international level, both spontaneously and on request, to effectively disrupt international financial flows \cite{Report2015}.

@@

The U.S. should request countries to proactively identify individuals and entities for inclusion in the UN Al Qaida Sanctions Committee list \cite{Report2015}.

@@

ISIL has shown itself prolific at recruiting foreign fighters, and the U.S. must try to stem this flow in its entirety, but also to prevent nuclear scientists from joining ISIL's causes.

@@

ISIL could attempt to purchase or steal a nuclear weapon, and so the nations which maintain a stockpile of nuclear weapons must remain under scrutiny. Nations of the highest interest that have nuclear weapons are Pakistan, North Korea, India, Turkey, Israel, and Russia. Israel has not been confirmed to have nuclear weapons, but it is widely believed that it has nuclear capabilites. Turkey is a NATO nuclear weapon sharing state and has U.S. tactical nuclear weapons stationed within its borders. However, it is unlikely that ISIL would be able to steal a nuclear weapon from either of these states, as those in Turkey are under U.S. protection and Israel has proven adept at hiding their arsenal. Of greatest concern is the possibility that any of these nuclear-capable states, North Korea in particular, would willingly sell a physics package to ISIL.

@@

The U.S. sees merits in pursuing both multilateral and bilateral effort in international strategic communications. Most of the times, the U.S. tends to favored intense bilateralism over multilateralism to achieve its objectives. However, following World War II, multilateral initiatives in the area of nuclear non-proliferation and disarmament increased in significance.  This multilateral transition was marked by the creation of the United Nations (UN) in 1945.  The UN, and several subsequent agencies, are the current approach taken by the U.S. to address nuclear proliferation concerns.
